Open Preprints in Ecology \& Evolution

### 2012-09-05 ms.tex
Added my byline and affiliation
New branch; brief outline of my section
This branch is for my description of arXiv.  I have inserted a comment
about the basic structure of the two or three paragraph passage.
a few small modifications

@@

Preprints servers such as arXiv are common in mathematics and physics, but
still a minority of papers in ecology and evolution use them. Preprints
servers allow author to make their manuscripts publicly available before, or
in parallel to, submitting them to journals for traditional peer-review. ...
The idea became popular with arXiv, and many physicists start their day with a
look at the new papers in arXiv \cite{gin11}

We will highlight advantages for both scientists and publishers.

The first and most often discussed advantage is speed \ref{fig:map}.

@@

%% @Tim: how can it improve the review process?

The review process as a whole is critically over-loaded, because the number of
active scientists increases, because the pressure to publish increases, and
because of an effect dubbed ``the tragedy of the reviewers commons'' REF.  In
the same times, rejection rates are high in most journals (REF), and when the
not invited to submit a revisions, authors are left with the impression that
they must start the whole process all over again. It's thus no surprize that
different initiatives emerged over the last few years, to decrease the time
spent in review. XXX et coll. (REF) called for the recycling and reuse of
peer-reviews: by attaching previous reviews, and detailed replies, to a new
submission, both the editor and the referees can jauge the work done on the
manuscript, and perhaps evaluate it with less prejudice. In a similar way, the
\emph{Peerage of Science} initiative allows authors to seek anonymous pre-
review by their peers. Some journals (LIST?) now accept to publish papers
which received good evaluations, effectively outsourcing the review process. A
widespread use of preprint servers can achieve the same goal of reducing the
time spent in review. By putting a manuscript out there for open comments and
criticisms, the authors will receive valuable feedback, and can improve the
version which will be submitted. With a rich enough community of scientists
depositing preprints, and commenting on them, the process of an open pre-
review can become widespread, and will overall increase the quality of first
submissions.
% Assuming that the quality of the paper is a good prior of its acceptance probability :-)

@@

Preprint servers also establish priority in a fair way. Some manuscript will
spend much more time in the review process. Public preprints server offer a
much fairer way to establish intellectual priority by making the work
available when done, and even if the exact organisation of the manuscript may
change. Surprisingly, there is perception in biology that public preprints
make it easier to steal ideas, as if scientific ideas only took form in
published material.
% TP This needs more development, I'm not sure I get your point
Mathematicians and physicists have embraced arXiv in part
to establish priority in a fair way\cite{cal12}.

Some of the responses to public preprints are surprising since they are,
essentially, the same as exchanging preprints among colleagues.
% TP Well, no. You trust your colleagues, and you trust that they will not give your
%    MS to anyone!
Prepublication reviews by a small network of colleagues is an important part
of the scientific process, which is attested by the fact that nearly all
published papers acknowledge comments by people not listed as co-authors.
Preprints servers simply offer a way to extend this network of colleagues to
the entire scientific community. It ensures that science is not constrained by
small networks of scientists exchanging ideas. Ginsparg made arXiv.org in part
for democratic reasons: he wanted everyone from graduate students in small
universities to Princeton professors to have access to the most recent
scientific \emph{ideas}.
% TP One important point to make: the MS and the idea are different things.
%    We should focus more on the later...
Ginsparg revolutionary idea was simply to use the power of the internet for
preprints, not just for the end product, so the process can be open from A to
Z, instead of being just open at the end of the process.
% TP Since when is the publishing buisness an open process? :-)

@@

While the practice is still rare, preprints are becoming more common in
biological sciences, which is experiencing faster growth in arXiv submissions
than any other fields \cite{cal12}. Also, most scientific journals are
preprint-friendly: Nature, PLOS, BMC, PNAS, Science (mostly)
\ref{table:policies}, and all the journals from Elsevier and Springer. Very
recently, the Ecological Society of America recently changed its policy to allow
public preprints (REF). In our field, few scientific publications will not
consider a manuscript submitted to arXiv. Still, many ecology \& evolution journals
adopt a ``by default'' hostile attitude towards preprints, mostly due to the
lack of clear policy of the publishers. As an example, Wiley-Blackwell, which
publishes some of the leading journal in the field, has no official policy on
the subject \ref{table:policies}.

\begin{table*}
    \centering
    \begin{tabular}{|ll|}
    \hline
    Publisher                                   & Policy \\
    \hline
    Springer                            	& Accept \\
    BMC                                 	& Accept \\
    Elsevier                            	& Accept \\
    Nature Publishing Group             	& Accept \\
    Public Library of Science           	& Accept \\
    Royal Society                       	& Accept \\
    National Academy of Science (USA)           & Accept \\
    Science                             	& Accept/Ambiguous \\
    Wiley-Blackwell                       	& No general policy \\
    Ecological Society of America       	& Refuse \\
    British Ecological Society                  & ? \\
    \hline
    \end{tabular}
    \caption{Policies for important publishers in ecology and evolution.}
    \label{table:policies}
\end{table*}

@@

Figshare (\href{http://figshare.com/}{http://figshare.com/})

@@

@@

% A short paragraph to conclude.

Responding to the rumour that they refused manuscripts submitted to arXiv,
Nature responded that ``Nature never wishes to stand in the way of communication
between researchers. We seek rather to add value for authors and the community
at large in our peer review, selection and editing'' \cite{nat05}.

>>>>

\section{Introduction}

Open preprints servers allow authors to make their manuscripts publicly
available before, or in parallel to, submitting them to journals for traditional
peer-review. The idea became popular with arXiv, an open preprint server, that
started 20 years ago \cite{gin11}. The idea is fundamentally simple: to make the
science available to the entire scientific community as soon as its done instead
of having to wait for the peer-review process. The point of arXiv and open
preprints servers is not to avoid the peer-review process. Almost all
manuscripts on arXiv are submitted to peer-review. The point is to open an
important phase of the publication cycle ...

In this article, we will first highlight the advantages of open preprints
servers for both scientists and publishers. We will also debunk a few
misconceptions, discuss the policies of major publishers in ecology an
evolution, and briefly review the most popular open preprint servers.

@@

The first and most often discussed advantage of arXiv and open preprints is
speed \ref{fig:map}. The time between submission and the official publication of
a manuscript can be measured in months, sometime in years. For all this time,
the research is only known to a select few: colleagues, editors, reviewers.
Thus, the science cannot be used, discussed, or reviewed by the wider scientific
community. ...

@@

% Tim's section:
The review process as a whole is critically over-loaded, because the number of
active scientists increases, because the pressure to publish increases, and
because of an effect dubbed ``the tragedy of the reviewers commons'' REF.  In
the same times, rejection rates are high in most journals (REF), and when the
not invited to submit a revisions, authors are left with the impression that
they must start the whole process all over again. It's thus no surprise that
different initiatives emerged over the last few years, to decrease the time
spent in review. XXX et coll. (REF) called for the recycling and reuse of
peer-reviews: by attaching previous reviews, and detailed replies, to a new
submission, both the editor and the referees can gauge the work done on the
manuscript, and perhaps evaluate it with less prejudice. In a similar way, the
\emph{Peerage of Science} initiative allows authors to seek anonymous pre-
review by their peers. Some journals (LIST?) now accept to publish papers
which received good evaluations, effectively outsourcing the review process. A
widespread use of preprint servers can achieve the same goal of reducing the
time spent in review. By putting a manuscript out there for open comments and
criticisms, the authors will receive valuable feedback, and can improve the
version which will be submitted. With a rich enough community of scientists
depositing preprints, and commenting on them, the process of an open pre-
review can become widespread, and will overall increase the quality of first
submissions.

@@

% @to myself: rewrite this paragraph:
Some of the responses to public preprints are surprising since they are,
essentially, the same as exchanging preprints among colleagues.
% TP Well, no. You trust your colleagues, and you trust that they will not give your
%    MS to anyone!
Prepublication reviews by a small network of colleagues is an important part
of the scientific process, which is attested by the fact that nearly all
published papers acknowledge comments by people not listed as co-authors.
Preprints servers simply offer a way to extend this network of colleagues to
the entire scientific community. It ensures that science is not constrained by
small networks of scientists exchanging ideas. Ginsparg made arXiv.org in part
for democratic reasons: he wanted everyone from graduate students in small
universities to Princeton professors to have access to the most recent
scientific \emph{ideas}.
% TP One important point to make: the MS and the idea are different things.
%    We should focus more on the later...
Ginsparg revolutionary idea was simply to use the power of the internet for
preprints, not just for the end product, so the process can be open from A to
Z, instead of being just open at the end of the process (and often not open at
all).
% TP Since when is the publishing buisness an open process? :-)
% PhDP: Wishful thinking :) Corrected...

Preprint servers also establish priority in a fair way. Since some manuscripts
will spend much more time in the review process, public preprints servers offer
a fairer way to establish intellectual priority by making the work available
when done. Surprisingly, there is perception in biology that public preprints
make it easier to steal ideas, while mathematicians and physicists have embraced
arXiv in part to establish intellectual priority in a fair way \cite{cal12}.

@@

While submitting to public preprints servers is still uncommon in ecology and
evolution, preprints are becoming more common in biological sciences. The
quantitative biology section in arXiv is experiencing faster growth in
submissions than any other fields \cite{cal12}. Also, most scientific journals
are preprint-friendly: Nature, PLOS, BMC, PNAS, Science (mostly)
\ref{table:policies}, and all the journals from Elsevier and Springer. Very
recently, the Ecological Society of America recently changed its policy to allow
public preprints (REF). In our field, few scientific publications will not
consider a manuscript submitted to arXiv.  Still, many ecology \& evolution
journals adopt a ``by default'' hostile attitude towards preprints, mostly due
to the lack of clear policy of the publishers. As an example, Wiley-Blackwell,
which publishes some of the leading journal in the field, has no official policy
on the subject \ref{table:policies}.

\begin{table*}
    \centering
    \begin{tabular}{|ll|}
    \hline
    Publisher                                   & Policy \\
    \hline
    Springer                            	& Accept \\
    BMC                                 	& Accept \\
    Elsevier                            	& Accept \\
    Nature Publishing Group             	& Accept \\
    Public Library of Science           	& Accept \\
    Royal Society                       	& Accept \\
    National Academy of Science (USA)           & Accept \\
    Ecological Society of America       	& Accept \\ % Victory!
    Science                             	& Accept/Ambiguous \\
    Wiley-Blackwell                       	& No general policy \\ % grrrrrr
    British Ecological Society                  & ? \\ % Waiting for their answer...
    % Oxford worth mentionning?
    \hline
    \end{tabular}
    \caption{Policies for important publishers in ecology and evolution.}
    \label{table:policies}
\end{table*}

@@

Figshare (\href{http://figshare.com/}{http://figshare.com/}) is an open servers
that allow scientists to submit any research output: manuscript, figures,
datasets, videos, theses, presentations, and so on. There are no rules to limit
what constitutes a research output: anything

@@

All content is licensed under the Creative Commons (CC-BY) license, except
datasets which are published under CC0.
% Small explanation

@@

% Poorly written:
Responding to the rumour that they refused manuscripts submitted to arXiv,
Nature responded that ``Nature never wishes to stand in the way of communication
between researchers. We seek rather to add value for authors and the community
at large in our peer review, selection and editing'' \cite{nat05}.

Open preprints server offer a great opportunity for open science, especially if
the community embrace the idea of discussing preprints. Initiatives like
Haldane's Sieve (\href{http://haldanessieve.org/}{http://haldanessieve.org/}), a
new blog discussing arXiv papers in population genetics, will help make arXiv
attractive for scientists looking to promote their work. These initiatives are
important to fully exploit the potential of open preprints servers.
